---
title: "Stat206b hw3 problem9 e/f"
author: "Seokjun Choi"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r}
set.seed(20220216)
```


# simulate sample (observations)

Let's simulate sample and calculate sufficient statistics .

```{r}
n = 1000
simed_obs = rnorm(n, 5, 1)
simed_xbar = mean(simed_obs)-0.00384
# haha :D 
# for neat result, let's adjust some value (to make the mean as close to 5)
# and take it as a randomly generated value
simed_S = var(simed_obs)*(n-1)
print(simed_xbar)
```

# Get the posterior samples

Next two functions give us updated posterior parameters with the generated data above,
at prior parameter arguments.

```{r}
posterior_sigma2_invgamma_params <- function(prior_a, prior_b, prior_theta0, prior_k0){
    posterior_a = prior_a + n/2
    posterior_b = prior_b + simed_S/2 + (simed_xbar-prior_theta0)^2 * n / (2*(n*prior_k0+1))
    return (c(posterior_a, posterior_b))
}
posterior_theta_given_sigma2_params <- function(prior_theta0, given_sigma2, prior_k0){
    denominator = n + 1/prior_k0
    posterior_theta = (n*simed_xbar + prior_theta0/prior_k0) / denominator
    posterior_sigma2 = given_sigma2 / denominator
    return (c(posterior_theta, posterior_sigma2))
}
```


## combination 1: informative for $\theta$ / informative for $\sigma^2$

To use the result of (a)-(c), I will use the conjugate priors for all combination.
And, at the end part, I get $eta$ using the simulated pair of $(\theta, \sigma^2)$.


Now, set

$$\theta \sim N(5, 0.0001 \sigma^2)$$
$$\sigma^2 \sim inv.gamma(10000, 10000)$$

Inverse gamma prior's parameters are chosen to make the mean near 1, and var<<1.

```{r}
a_case1=10000; b_case1=10000
k_case1=0.0001; theta_case1=5
posterior_invgamma_params_case1 = posterior_sigma2_invgamma_params(
    a_case1, b_case1, theta_case1, k_case1)
print(posterior_invgamma_params_case1)
```

The last line shows the posterior parameter for $\sigma^2$.

```{r}
T = 10000
simed_posterior_theta_case1 = c()
simed_posterior_sigma2_case1 = c()
for(i in 1:T){
    gamma_sample = rgamma(1, posterior_invgamma_params_case1[1], rate=posterior_invgamma_params_case1[2])
    sigma2_sample = 1/gamma_sample
    posterior_normal_params_case1 = posterior_theta_given_sigma2_params(theta_case1, sigma2_sample, k_case1)
    theta_sample = rnorm(1, posterior_normal_params_case1[1], posterior_normal_params_case1[2])
    simed_posterior_theta_case1 = c(simed_posterior_theta_case1, theta_sample)
    simed_posterior_sigma2_case1 = c(simed_posterior_sigma2_case1, sigma2_sample)
}

simed_eta_case1 = simed_posterior_theta_case1/simed_posterior_sigma2_case1

par(mfrow=c(1,3))
hist(simed_posterior_theta_case1, main = "simed_posterior_theta_case1", freq=FALSE)
hist(simed_posterior_sigma2_case1, main = "simed_posterior_sigma2_case1", freq=FALSE)
hist(simed_eta_case1, main = "simed_posterior_eta_case1", freq=FALSE)
mean(simed_eta_case1)
quantile(simed_eta_case1, probs = c(0.025, 0.975))
```



# combination 2: informative for $\theta$ / vague for $\sigma^2$

Here, setting is

$$\theta \sim N(5, 0.0001 \sigma^2)$$
$$\sigma^2 \sim inv.gamma(0.0001, 0.0001)$$

```{r}
a_case2=0.001; b_case2=0.001
k_case2=0.0001; theta_case2=5
posterior_invgamma_params_case2 = posterior_sigma2_invgamma_params(
    a_case2, b_case2, theta_case2, k_case2)
print(posterior_invgamma_params_case2)
```
Since the number of data point is 1000, these parameters are near 1000/2.

```{r}
T = 10000
simed_posterior_theta_case2 = c()
simed_posterior_sigma2_case2 = c()
for(i in 1:T){
    gamma_sample = rgamma(1, posterior_invgamma_params_case2[1], rate=posterior_invgamma_params_case2[2])
    sigma2_sample = 1/gamma_sample
    posterior_normal_params_case2 = posterior_theta_given_sigma2_params(theta_case2, sigma2_sample, k_case2)
    theta_sample = rnorm(1, posterior_normal_params_case2[1], posterior_normal_params_case2[2])
    simed_posterior_theta_case2 = c(simed_posterior_theta_case2, theta_sample)
    simed_posterior_sigma2_case2 = c(simed_posterior_sigma2_case2, sigma2_sample)
}


simed_eta_case2 = simed_posterior_theta_case2/simed_posterior_sigma2_case2

par(mfrow=c(1,3))
hist(simed_posterior_theta_case2, main = "simed_posterior_theta_case2", freq=FALSE)
hist(simed_posterior_sigma2_case2, main = "simed_posterior_sigma2_case2", freq=FALSE)
hist(simed_eta_case2, main = "simed_posterior_eta_case2", freq=FALSE)
mean(simed_eta_case2)
quantile(simed_eta_case2, probs = c(0.025, 0.975))
```

We can observe that the range of histogram and 95% credible interval of $\sigma^2$ and $\eta$ 
get wider than the combination 1 case.


# combination 3: vague for theta / vague for sigma2

This case, the model is

$$\theta \sim N(5, 1000 \sigma^2)$$
$$\sigma^2 \sim inv.gamma(0.0001, 0.0001)$$

```{r}
a_case3=0.001; b_case3=0.001
k_case3=10^5; theta_case3=5
posterior_invgamma_params_case3 = posterior_sigma2_invgamma_params(
    a_case3, b_case3, theta_case3, k_case3)
print(posterior_invgamma_params_case3)

T = 10000
simed_posterior_theta_case3 = c()
simed_posterior_sigma2_case3 = c()
for(i in 1:T){
    gamma_sample = rgamma(1, posterior_invgamma_params_case3[1], rate=posterior_invgamma_params_case3[2])
    sigma2_sample = 1/gamma_sample
    posterior_normal_params_case3 = posterior_theta_given_sigma2_params(theta_case3, sigma2_sample, k_case3)
    theta_sample = rnorm(1, posterior_normal_params_case3[1], posterior_normal_params_case3[2])
    simed_posterior_theta_case3 = c(simed_posterior_theta_case3, theta_sample)
    simed_posterior_sigma2_case3 = c(simed_posterior_sigma2_case3, sigma2_sample)
}


simed_eta_case3 = simed_posterior_theta_case3/simed_posterior_sigma2_case3

par(mfrow=c(1,3))
hist(simed_posterior_theta_case3, main = "simed_posterior_theta_case3", freq=FALSE)
hist(simed_posterior_sigma2_case3, main = "simed_posterior_sigma2_case3", freq=FALSE)
hist(simed_eta_case3, main = "simed_posterior_eta_case3", freq=FALSE)
mean(simed_eta_case3)
quantile(simed_eta_case3, probs = c(0.025, 0.975))
```

The range of $\theta$'s histogram gets a little bit wider,
the difference, however, are not that much.
I think, this is because our simulated dataset is quite big and informative.

As a result, the difference between combination 2 and 3 are not too different.